An efficient acyclic contact planner for multiped robots

### 2017-01-12 body/appendix_algo.tex
[Appendix] + starting results

\section{Derivation of the manipulability ellipsoid}
\label{app:manipulability}

Again, we assume that $\mathbf{J}$ is full rank. We discard the $k$ indices, and write the pseudo-inverse of $\mathbf{J}$ as $\mathbf{J}^{\dagger}$.
\begin{eqnarray*}
\mathbf{\dot{p}} & =  & \mathbf{J} \dot{\mathbf{q}} \\ 
\mathbf{\dot{q}} & =  & \mathbf{J}^{\dagger} \dot{\mathbf{p}} \\ 
\mathbf{\dot{q}}^T & =  & \dot{\mathbf{p}}^T \mathbf{J}^{\dagger T} \\ 
\mathbf{\dot{q}}^T\mathbf{\dot{q}} & =  & \dot{\mathbf{p}}^T \mathbf{J}^{\dagger T} \mathbf{J}^{\dagger} \dot{\mathbf{p}}\\ 
\end{eqnarray*}

Then, proving that $\mathbf{J}^{\dagger T} \mathbf{J}^{\dagger} = (\mathbf{J}\mathbf{J}^T)^{-1}$ is straight-forward,
using the SVD decomposition of each term.

>>>>

%~ \section{Derivation of the manipulability ellipsoid}
%~ \label{app:manipulability}
%~ 
%~ Again, we assume that $\mathbf{J}$ is full rank. We discard the $k$ indices, and write the pseudo-inverse of $\mathbf{J}$ as $\mathbf{J}^{\dagger}$.
%~ \begin{eqnarray*}
%~ \mathbf{\dot{p}} & =  & \mathbf{J} \dot{\mathbf{q}} \\ 
%~ \mathbf{\dot{q}} & =  & \mathbf{J}^{\dagger} \dot{\mathbf{p}} \\ 
%~ \mathbf{\dot{q}}^T & =  & \dot{\mathbf{p}}^T \mathbf{J}^{\dagger T} \\ 
%~ \mathbf{\dot{q}}^T\mathbf{\dot{q}} & =  & \dot{\mathbf{p}}^T \mathbf{J}^{\dagger T} \mathbf{J}^{\dagger} \dot{\mathbf{p}}\\ 
%~ \end{eqnarray*}
%~ 
%~ Then, the equality $\mathbf{J}^{\dagger T} \mathbf{J}^{\dagger} = (\mathbf{J}\mathbf{J}^T)^{-1}$ follows from the SVD decomposition of each term
%~ ~\citep{ben2003generalized}.

### 2017-01-12 body/appendix_roms.tex
merging appendices
[Appendix] + starting results

\section{Generating the $W$ volumes for HRP-2}
\label{app:rom}

In this Appendix,
we detail our method to generate the volumes $W$ used
in RB-RRT, with the example of HRP-2.

We consider the robot HRP-2, and proceed to the decomposition 
of its kinematic chain into four limbs $R^k$.
The arms are connected to the shoulders, and the legs to the root.
The obtained volumes $W$ are shown in Figure~\ref{fig:hrp2_w}.

@@

\begin{figure}
\centering
  \begin{overpic}[width=1\linewidth]{figures/roms}
	\end{overpic}
\caption{Different approximations of the range of motion of the right arm HRP-2. Left: non convex-hull, computed with the powercrust algorithm~\citep{Amenta:2001:PC:376957.376986}. Middle:
convex hull of the reachable workspace. Right: Simplified hull used in our experiments.}
		   \label{fig:hrp2_roms}
\end{figure}

Figure~\ref{fig:hrp2_roms} illustrate the obtained $W^k$ for HRP-2.
Regarding the procedure, we can see that step 2 is conservative (Figure~\ref{fig:hrp2_roms}--right), which 
is acceptable, especially because the lost set essentially relates to configurations close to singularity (they are close to the boundaries of the reachable workspace). We choose again to be less complete
but more efficient, regarding the number of collision tests to be performed by RB-RRT.
In step 1 on the other hand, selecting the convex hull (Figure~\ref{fig:hrp2_roms}--middle) instead of a minimum encompassing shape (Figure~\ref{fig:hrp2_roms}--left) may introduce false positives.
Concretely, because the false positive set intersects with $W^0$, the scaling volume of the robot torso, the induced error is compensated,
as verified by the results shown by Table~\ref{tab:requestpercent}.

>>>>

%~ \clearpage
\section{Generating the $W$ volumes for HRP-2}
\label{app:rom}

We detail our method to generate the volumes $W$ used
in RB-RRT, with the example of HRP-2.
The kinematic chain is split into four limbs $R^k$.
The arms are connected to the shoulders, and the legs to the root.
The obtained volumes $W$ are shown in Figure~\ref{fig:hrp2_w}.

@@

\begin{figure}
\centering
  \begin{overpic}[width=1\linewidth]{figures/roms}
	\end{overpic}
\caption{Different approximations of the range of motion of the right arm of HRP-2. Left: non convex-hull, computed with the powercrust algorithm~\citep{Amenta:2001:PC:376957.376986}. Middle:
convex hull of the reachable workspace. Right: Simplified hull used in our experiments.}
		   \label{fig:hrp2_roms}
\end{figure}

Figure~\ref{fig:hrp2_roms} illustrate the obtained $W^k$ for HRP-2.
Regarding the procedure, we can see that step 2 is conservative (Figure~\ref{fig:hrp2_roms}--right), which 
is acceptable, especially because the lost set essentially relates to configurations close to singularity (they are close to the boundaries of the reachable workspace, and
often not \gls{contact reachable}, as illustrated in Figure~\ref{fig:dedefeas}, where the exterior boundaries of the reachable workspace appear
red, thus not belonging to $C_{Contact}^0$). We choose again to be less complete but more efficient, regarding the number of collision tests to be performed by RB-RRT.
In step 1 on the other hand, selecting the convex hull (Figure~\ref{fig:hrp2_roms}--middle) instead of a minimum encompassing shape (Figure~\ref{fig:hrp2_roms}--left) may introduce false positives.
Concretely, because the false positive set intersects with $W^0$, the scaling volume of the robot torso, the induced error is compensated,
as verified by the results shown by Table~\ref{tab:requestpercent}.

### 2017-01-12 body/efort.tex
[Appendix] + starting results

% !TEX root =  ../main.tex
\subsection{Manipulability-based heuristics for contact selection}
This Section proposes heuristics to select a contact that optimizes desiered capabilities.
%~ These heuristics allow to locally optimize a contact location with respect to a particular task.
For instance, one can be interested in configurations that allow to efficiently exert a force in the global direction of motion,
a high velocity in a given direction, or to stay away from singular configurations.
In this section, we derive three such heuristics from the work on manipulability by \cite{Yoshikawa1984}, that we present first.

\subsubsection{The force and velocity ellipsoids:}
For a limb configuration $\mathbf{q}^k$, the Jacobian matrix
 $\mathbf{J}^k(\mathbf{q}^k)$   defines the relation:

\begin{equation} \label{jac}
\mathbf{\dot{p}}^k = \mathbf{J}^k(\mathbf{q}^k) \dot{\mathbf{q}}^k
\end{equation}

For clarity in the rest of the section we omit the $k$ indices and write $\mathbf{J}^k(\mathbf{q}^k)$ as $\mathbf{J}$.

As a linear approximation of a forward-kinematics function, $\mathbf{J}$ describes how small
variations from the configuration $\mathbf{q}$ affect the position vector $\mathbf{p}$.
%~ It can be seen as the transformation mapping a velocity in the configuration space into a velocity in
%~ the Euclidian space.

Now we consider the unit ball in the configuration space $C$ defined by the set of joint velocities
for which the norm is at most 1:
\begin{equation} \label{ball}
||\dot{\mathbf{q}}||^2 \leq 1
\end{equation}

We assume that $\mathbf{J}$ is full rank (we are not interested in singular configurations, which we discard).
From \eqref{jac} we can thus obtain the following equality (Appendix~\ref{app:manipulability}):
\begin{equation} \label{jaceq}
\mathbf{\dot{p}}^T(\mathbf{J}\mathbf{J}^T)^{-1}\mathbf{\dot{p}} = \dot{\mathbf{q}}^T \dot{\mathbf{q}}
\end{equation}

We can use \eqref{jaceq} to map the ball into an ellipsoid in the Euclidian space $\mathbb{R}^m$:

\begin{equation} \label{ellipsoid}
\mathbf{\dot{p}}^T(\mathbf{J}\mathbf{J}^T)^{-1}\mathbf{\dot{p}} \leq 1
\end{equation}
This ellipsoid is called the manipulability ellipsoid, or velocity ellipsoid, introduced by \cite{Yoshikawa1984}. It describes the set of end-effector velocities that can
be reached under the constraint \eqref{ball} for the current configuration.
The longer the axis of the ellipsoid is, the faster the end-effector can move along the direction of the axis.
Figure~\ref{sec:efort_ellipsoid} - left shows the velocity ellipsoid for different configurations of a manipulator with two degrees of freedom.

\begin{figure}[!tbp]
  \centering
	\begin{overpic}[width=1\linewidth]{figures/EFORT/ellipsoid}
		\put (8.5,1.2) {\small{Velocity ellipsoid}}
		\put (62.5,1.2) {\small{Force ellipsoid} \tiny{(scale 0.5)}}
	\end{overpic}
  % \includegraphics[width=1\linewidth]{ressources/screens/sampling/canap/bd2_cmyk}
  \caption{Examples of velocity and force ellipsoids for a manipulator composed of 2 dofs and 2 segments.
Only the horizontal and vertical speeds are shown (not the rotation speeds), since it would require being able to draw in four dimensions.}
		   \label{sec:efort_ellipsoid}
\end{figure}

Similarly to the velocity ellipsoid, \citeauthor{Yoshikawa1984} also defines the force ellipsoid.
Considering: a force vector $\mathbf{f}$ expressed in the task space $\mathbb{R}^m$;
the equivalent joint torque vector $\bm{\tau}$;
we can define the mechanical work in both spaces:
\begin{equation*} \label{power}
\dot{\mathbf{q}}^T \bm{\tau} = \dot{\mathbf{p}}^T \mathbf{f}
\end{equation*}

Exploiting \eqref{jac}, we can easily see that the set of achievable forces in $\mathbb{R}^m$ subject to the constraint:
\begin{equation*} \label{ballforce}
||\bm{\tau}||^2 \leq 1
\end{equation*}
is the so-called force ellipsoid (Figure~\ref{sec:efort_ellipsoid} - right):
\begin{equation} \label{ellipsoidforce}
\mathbf{f}^T (\mathbf{J}\mathbf{J}^T) \mathbf{f} \leq 1
\end{equation}

@@

In our problem, to compare candidate configurations, we include the quality of the contact surface, and choose $\mathbf{m}$ as the direction
opposite to the local motion (thus given by the difference between two consecutive root positions):

@@

\begin{equation} \label{ellipsoid}
h_{w}(\mathbf{q}) = \sqrt{det(\mathbf{J}\mathbf{J}^T)}
\end{equation}
$h_{w}$ measures the ``distance'' of a given configuration to singularity. When $h_{w}$ is equal to 0, the configuration is singular;
the greater $h_{w}$ is, the further away the configuration is from singularity.
%~ We then derive:
%~ \begin{equation}
%~ h_{w}(\mathbf{q}, \mathbf{m}) = w(\mathbf{q})  (\mu \mathbf{n}^T \mathbf{m})
%~ \end{equation}

$h_{\textrm{\it EFORT}}$, $h_{vel}$ and $h_{w}$ define three kinematic heuristics, fast to compute (indeed, thanks to our sampling-based approach, the Jacobian and inverse products
can be precomputed off-line), that allow the planner to select the best
candidates according to user-defined criterion.
%~ This results in faster computation times, and a reduced number of overall contact switches.

>>>>

% !TEX root =  ../main.tex
\subsection{Manipulability-based heuristics for contact selection}
This Section proposes heuristics to select a contact that optimizes desired capabilities.
%~ These heuristics allow to locally optimize a contact location with respect to a particular task.
For instance, one can be interested in configurations that allow to efficiently exert a force in the global direction of motion,
a high velocity in a given direction, or to stay away from singular configurations.
We derive three such heuristics from the work by \cite{Yoshikawa1984}, recalled here. %, that we present first.

\subsubsection{The force and velocity ellipsoids:}

\begin{figure}[!tbp]
  \centering
	\begin{overpic}[width=1\linewidth]{figures/EFORT/ellipsoid}
		\put (8.5,1.2) {\small{Velocity ellipsoid}}
		\put (62.5,1.2) {\small{Force ellipsoid} \tiny{(scale 0.5)}}
	\end{overpic}
  \caption{Examples of velocity and force ellipsoids for a manipulator composed of 2 dofs and 2 segments.
Only the horizontal and vertical speeds are shown (not the rotation speeds).}
 %~ since it would require being able to draw in four dimensions.}
		   \label{sec:efort_ellipsoid}
\end{figure}

We consider: a limb configuration $\mathbf{q}^k$; its end effector position $\mathbf{p}^k$; its Jacobian matrix
$\mathbf{J}^k(\mathbf{q}^k)$; a force $\mathbf{f}$ exerted by the end effector. For clarity in the rest of the section we omit the $k$ indices and write $\mathbf{J}^k(\mathbf{q}^k)$ as $\mathbf{J}$.
\citeauthor{Yoshikawa1984} defines the velocity~\ref{eq:vel} and force~\ref{eq:vel} ellipsoids:

 \begin{equation} 
 \label{eq:vel}
\mathbf{\dot{p}}^T(\mathbf{J}\mathbf{J}^T)^{-1}\mathbf{\dot{p}} \leq 1 
\end{equation}

 \begin{equation} 
 \label{eq:for}
\mathbf{f}^T (\mathbf{J}\mathbf{J}^T) \mathbf{f} \leq 1
\end{equation}

They describe the set of end-effector velocities (respectively forces) that can
be reached under the constraint $||\dot{\mathbf{q}}||^2 \leq 1$ for the current configuration.
The longer the axis of the ellipsoid is, the more important the velocity (resp. force) of the end-effector the direction of the axis can be (Figure~\ref{sec:efort_ellipsoid}).

%~ Similarly, \citeauthor{Yoshikawa1984} also defines the force ellipsoid (Figure~\ref{sec:efort_ellipsoid} - right):

%~ \begin{equation} \label{ellipsoidforce}
%~ \mathbf{f}^T (\mathbf{J}\mathbf{J}^T) \mathbf{f} \leq 1
%~ \end{equation}
%~ with $\mathbf{f}$ is a force vector expressed in the task space $\mathbb{R}^m$;

@@

In our problem, to compare candidate configurations, we include the quality of the contact surface, and choose $\mathbf{m}$ as the direction
opposite to the local motion (given by the difference between two consecutive root positions):

@@

\begin{equation} \label{ellipsoid}
h_{w}(\mathbf{q}) = \sqrt{det(\mathbf{J}\mathbf{J}^T)}
\end{equation}
$h_{w}$ measures the ``distance'' of a given configuration to singularity. When $h_{w}$ is equal to 0, the configuration is singular;
the greater $h_{w}$ is, the further away the configuration is from singularity.

%~ $h_{\textrm{\it EFORT}}$, $h_{vel}$ and $h_{w}$ define three kinematic heuristics, fast to compute (indeed, thanks to our sampling-based approach, the Jacobian and inverse products
%~ can be precomputed off-line), that allow the planner to select the best
%~ candidates according to user-defined criterion.

%~ This results in faster computation times, and a reduced number of overall contact switches.

%~ where the $\dot$ symbol represents the time derivative.  It describes the set of end-effector velocities that can
%~ be reached under the constraint \eqref{ball} for the current configuration.
%~ The longer the axis of the ellipsoid is, the faster the end-effector can move along the direction of the axis (Figure~\ref{sec:efort_ellipsoid}).
 %~ 
%~ Similarly, \citeauthor{Yoshikawa1984} also defines the force ellipsoid(Figure~\ref{sec:efort_ellipsoid} - right):
 %~ 
%~ \begin{equation} \label{ellipsoidforce}
%~ \mathbf{f}^T (\mathbf{J}\mathbf{J}^T) \mathbf{f} \leq 1
%~ \end{equation}
%~ where $\mathbf{f}$ is a force vector $\mathbf{f}$ expressed in the task space $\mathbb{R}^m$;
%~ the equivalent joint torque vector $\bm{\tau}$;
 %~ 
%~ \begin{equation} \label{jac}
%~ \mathbf{\dot{p}}^k = \mathbf{J}^k(\mathbf{q}^k) \dot{\mathbf{q}}^k
%~ \end{equation}
%~ 
%~ For clarity in the rest of the section we omit the $k$ indices and write $\mathbf{J}^k(\mathbf{q}^k)$ as $\mathbf{J}$.
 %~ 
%~ As a linear approximation of a forward-kinematics function, $\mathbf{J}$ describes how small
%~ variations from the configuration $\mathbf{q}$ affect the position vector $\mathbf{p}$.

%~ Now we consider the unit ball in the configuration space $C$ defined by the set of joint velocities
%~ for which the norm is at most 1:
%~ \begin{equation} \label{ball}
%~ ||\dot{\mathbf{q}}||^2 \leq 1
%~ \end{equation}
%~ 
%~ We assume that $\mathbf{J}$ is full rank (we are not interested in singular configurations, which we discard).
%~ From \eqref{jac} we can thus obtain the following equality (Appendix~\ref{app:manipulability}):
%~ \begin{equation} \label{jaceq}
%~ \mathbf{\dot{p}}^T(\mathbf{J}\mathbf{J}^T)^{-1}\mathbf{\dot{p}} = \dot{\mathbf{q}}^T \dot{\mathbf{q}}
%~ \end{equation}

%~ We can use \eqref{jaceq} to map the ball into an ellipsoid in the Euclidian space $\mathbb{R}^m$:

%~ \begin{equation} \label{ellipsoid}
%~ \mathbf{\dot{p}}^T(\mathbf{J}\mathbf{J}^T)^{-1}\mathbf{\dot{p}} \leq 1
%~ \end{equation}
%~ This ellipsoid is called the manipulability ellipsoid, or velocity ellipsoid, introduced by \cite{Yoshikawa1984}. It describes the set of end-effector velocities that can
%~ be reached under the constraint \eqref{ball} for the current configuration.
%~ The longer the axis of the ellipsoid is, the faster the end-effector can move along the direction of the axis.
%~ Figure~\ref{sec:efort_ellipsoid} - left shows the velocity ellipsoid for different configurations of a manipulator with two degrees of freedom.

%~ \begin{figure}[!tbp]
  %~ \centering
	%~ \begin{overpic}[width=1\linewidth]{figures/EFORT/ellipsoid}
		%~ \put (8.5,1.2) {\small{Velocity ellipsoid}}
		%~ \put (62.5,1.2) {\small{Force ellipsoid} \tiny{(scale 0.5)}}
	%~ \end{overpic}
  %~ \caption{Examples of velocity and force ellipsoids for a manipulator composed of 2 dofs and 2 segments.
%~ Only the horizontal and vertical speeds are shown (not the rotation speeds), since it would require being able to draw in four dimensions.}
		   %~ \label{sec:efort_ellipsoid}
%~ \end{figure}

%~ Similarly to the velocity ellipsoid, \citeauthor{Yoshikawa1984} also defines the force ellipsoid.
%~ Considering: a force vector $\mathbf{f}$ expressed in the task space $\mathbb{R}^m$;
%~ the equivalent joint torque vector $\bm{\tau}$;
%~ we can define the mechanical work in both spaces:
%~ \begin{equation*} \label{power}
%~ \dot{\mathbf{q}}^T \bm{\tau} = \dot{\mathbf{p}}^T \mathbf{f}
%~ \end{equation*}
%~ 
%~ 
%~ Exploiting \eqref{jac}, we can easily see that the set of achievable forces in $\mathbb{R}^m$ subject to the constraint:
%~ \begin{equation*} \label{ballforce}
%~ ||\bm{\tau}||^2 \leq 1
%~ \end{equation*}
%~ is the so-called force ellipsoid (Figure~\ref{sec:efort_ellipsoid} - right):
%~ \begin{equation} \label{ellipsoidforce}
%~ \mathbf{f}^T (\mathbf{J}\mathbf{J}^T) \mathbf{f} \leq 1
%~ \end{equation}

### 2017-01-12 body/equilibrium.tex
ongoing intro rewriting
Move Section about heuristics to Appendix
[Appendix] + starting results

% !TEX root =  ../main.tex
\section{Heuristics for contact selection}
\label{sec:heuristics}
\subsection{A heuristic for robust static equilibrium}
The planner is designed so that any generated contact configuration is in static equilibrium.
%~ Equilibrium is of course critical in multipedal locomotion. For this reason 
We are interested in a robust 
criterion, that ensures that the robot remains in equilibrium in a real-world application, regardless of perception and control uncertainties.

@@

\subsubsection{Conditions for static equilibrium:}
We first define the variables of the problem, for $e$ contact points, expressed in world coordinates:
\begin{itemize}
\item $\mathbf{c} \in \mathbb{R}^3$ is the robot center of mass (COM);
%~ \item $\mathbf{L}  \in \mathbb{R}^3 $ is the angular momentum at the COM;
\item $\mathbf{m} \in \mathbb{R}$ is the robot mass;
\item $\mathbf{g} = [0,0,-9.81]^T$ is the gravity acceleration;
\item $\mu$ is the friction coefficient;
\item for the i-th contact point $1 \leq i \leq e$:
	\begin{itemize}
	\item $\mathbf{p_i}$ is the contact position;
	\item $\mathbf{f_i}$ is the force applied at $\mathbf{p_i}$;
	\item $\mathbf{n}_i,\mathbf{t}_{i1},\mathbf{t}_{i2}$ form a local Cartesian coordinate system centered at $\mathbf{p_i}$. $\mathbf{n}_i$ is aligned
	with the contact surface normal, and the $\mathbf{t}_i$s are tangent vectors.
	\end{itemize}
\end{itemize}

>>>>

% !TEX root =  ../main.tex
\section{Heuristics and criteria for contact selection}
\label{sec:heuristics}
\subsection{A criterion for robust static equilibrium}
The planner is designed so that any generated contact configuration is in static equilibrium.
%~ Equilibrium is of course critical in legged locomotion. For this reason 
We are interested in a robust 
criterion, that ensures that the robot remains in equilibrium in a real-world application, regardless of perception and control uncertainties.

@@

\subsubsection{Conditions for static equilibrium:}
We first define the variables of the problem, for $e$ contact points, expressed in world coordinates:
\begin{itemize}
\item $\mathbf{c} \in \mathbb{R}^3$ is the robot center of mass (COM);
%~ \item $\mathbf{L}  \in \mathbb{R}^3 $ is the angular momentum at the COM;
\item $m \in \mathbb{R}$ is the robot mass;
\item $\mathbf{g} = [0,0,-9.81]^T$ is the gravity acceleration;
\item $\mu$ is the friction coefficient;
\item for the i-th contact point $1 \leq i \leq e$:
	\begin{itemize}
	\item $\mathbf{p_i}$ is the contact position;
	\item $\mathbf{f_i}$ is the force applied at $\mathbf{p_i}$;
	\item $\mathbf{n}_i,\mathbf{t}_{i1},\mathbf{t}_{i2}$ form a local Cartesian coordinate system centered at $\mathbf{p_i}$. $\mathbf{n}_i$ is aligned
	with the contact surface normal, and the $\mathbf{t}_i$s are tangent vectors.
	\end{itemize}
\end{itemize}

### 2017-01-12 body/results.tex
[Appendix] + starting results

% !TEX root =  ../main.tex
\section{Results}
\label{sec:results}
In this Section we present some of the results obtained with our planner. The complete sequences computed are shown in the companion video.
Specifically, we demonstrate the planner for two really different robots, in a large variety of environments: the humanoid HRP-2 and the quadruped HyQ.
%For each scenario we indicate the chosen heuristics. 
We also provide a performance analysis, which shows that the planner is compatible with \gls{interactive} applications,
and present the success rates obtained in each scenario. Moreover, we demonstrate the interest of our robustness criterion in the different computed poses.
Finally, a last example suggests possible applications to dexterous manipulation.

In our previous work~\citep{tonneauisrr15} additional results are demonstrated with various virtual avatars (Figure~\ref{fig:robots_old}).
In this extension we choose to focus on actual robots. We invite the interested reader to watch the ISRR video (\url{http://youtu.be/LmLAHgGQJGA}), and 
to refer to the previous paper for a discussion on these results.

>>>>

% !TEX root =  ../main.tex
\section{Results}
\label{sec:results}
In this Section we present some of the results obtained with our planner. The complete sequences computed are shown in the companion video.
Specifically, we demonstrate the planner for two really different robots, in a large variety of environments: the humanoid HRP-2 and the quadruped HyQ.
Finally, a last example suggests possible applications to dexterous manipulation.
At the end of the video, we validate our contact plans by exhibiting a solution to the interpolation problem $\mathcal{P}_3$ for these scenarios.

At the end of this Section, we provide the \textit{interactive} computation times obtained in each case.
We also compare the times obtained with HRP-2 with respect to previous demonstrations.

These scenarios complete the ones demonstrated on virtual avatars (Figure~\ref{fig:robots_old}) in our previous ISRR paper~\citep{tonneauisrr15} and video\footnote{\url{http://youtu.be/LmLAHgGQJGA}}.
%~ (Figure~\ref{fig:robots_old}).
 %~ We invite the interested reader to watch the ISRR video (\url{http://youtu.be/LmLAHgGQJGA}), and 
%~ to refer to the previous paper for a discussion on these results.
